---
title: Quotes
author: Charlotte Wickham
date: May 8th 2018
output: github_document
always_allow_html: yes
---

```{r setup, include = FALSE}
library(printr)
```


```{r, message=FALSE}
library(tidyverse)
library(lobstr)
library(rlang)
```

# Part I: Quotes

When do you need to surround something in quote marks?

## Quiz

What's the difference between
```{r, results="hide"}
3
```
and 
```{r, results="hide"}
"3"
```
?

*(For experienced users, how is this one different?)*
```{r, eval = FALSE}
`3`
```

## Quotes tell R something is a character string

`"` and `'` are used in R to form character strings.  

`3` is the number three,  
`"3"` is the character string containing the digit 3.
```{r}
str(3)
str("3")
```

## Backticks are for *non-syntatic* names

Backticks `` ` `` are used for a object **name** (also known as symbol).  You don't need them unless the name is unusual and wouldn't usually be considered a valid R object name, e.g. it begins with a number, contains spaces, or other reserved characters.

Here, we are asking for an object called `3`,
```{r, error = TRUE}
`3`
```
but one doesn't exist, so R returns an error.  If we really wanted to confuse ourselves, we could create an object with the name 3:

```{r}
`3` <- 4
`3`
```
But, I wouldn't advise it.  More often, weird names occur as the result of inputting data where column names might involve spaces or special characters.

```{r, eval = FALSE}
?Quotes          # Read about ",' and `
?make.names      # Read rules for names to be valid
```

## Quiz 

```{r}
c <- 2
```

In the code below: 

* Which `c` is a *character string*?
* Which `c` refers to a *the name of an object*?
* Which `c` refers to a *function name*?
* Which `c` refers to an *argument name*?
  
```{r, results="hide"}
c(c = "c", c)
```

## Quiz: solutions

The character string containing the letter `c`
```
c(c = "c", c)
       ^
       |
```

An object with the name `c`
```
c(c = "c", c)
           ^
           |
```

Function name
```
c(c = "c", c)
^
|
```

Argument name
```
c(c = "c", c)
  ^
  |
```


## You don't need quotes around function names or argument names

Function names and argument names **very rarely** need quotes.  If you do put quotes around them, you probably won't get an error, but it's considered bad style.

These all work, and give the same result:
```{r, results="hide"}
"mean"( x  =  c (1:5))
"mean"("x" =  c (1:5))
"mean"("x" = "c"(1:5))
 mean ( x  = "c"(1:5))
 mean ("x" = "c"(1:5))
```

But, you should write:
```{r, results = "hide"}
mean(x = c(1:5))
```

## Except...if you need a backtick `` ` ``

The exception is function names or argument names that aren't considered valid R names, and you'll need backticks (`` ` ``).

E.g. the function with the name `+`
```{r}
`+`(1, 2)
```

OK that leaves **argument values**... to quote or not to quote?

## Strategy #0: Trial and `Error` + Practice

Where most of us start!  

The more you use a function the better you get at guessing/remembering what it is expecting.

But, it helps to have strategies to remove the guessing.

## Strategy #1: What are you referring to?

| You are referring to:                                |            |
|------------------------------------------------------|------------|
| an **object** that exists in your workspace by name  | No quotes  |
| a particular character **string**                    | Quotes     |

```{r}
n_times <- 10
```

I want to repeat the string "A", `n_times` times:
```{r}
rep(x = "A", times = n_times)
```

`A` is the **character string** I want to repeat, so it needs quotes, `n_times` is the **name of an object**, so it doesn't need quotes.  

If I stored the string I wanted to repeat in `str_to_repeat`:
```{r}
string_to_repeat <- "A"
```

Now, my `x` argument is the **name of an object** so it doesn't need quotes:
```{r}
rep(x = string_to_repeat, times = n_times)
```


## Strategy #1: Your Turn

```{r}
books <- c("The Art of R Programming", "R for Data Science",
  "Advanced R")
R <- R.version.string
```

I want to see where the character string `R` occurs in the titles in my `books` variable.

**Which code is correct?**

```{r, eval = FALSE}
str_view(string = "books", pattern = "R")
str_view(string = "books", pattern =  R )
str_view(string =  books,  pattern = "R")
str_view(string =  books,  pattern =  R )
```

## Strategy #1: Your Turn Solution

`books` is the name of an object, it doesn't need quotes, 
`R` is a specific character string, it needs quotes, so the correct option is:
```{r}
str_view(string = books, pattern = "R")
```
(Caveat: this strategy may not work if the function **quotes** its arguments, more in Quotation)

## Strategy #2: Read the documentation

If the argument value is described as:

* a character string, or string, use quotes,
* an object, name, symbol, or expression don't use quotes

## Strategy #2: Example

The `pull()` function in dplyr extracts a column from a data frame.  

If I want the `cyl` column from `mtcars`, do I want `pull(mtcars, cyl)` or `pull(mtcars, "cyl")`?

```{r tidy=FALSE, printr.help.sections=c('arguments'), comment=''}
?dplyr::pull
```

No quotes needed since this is a name.  Also, look at examples section and see:

```{r}
mtcars %>% pull(cyl)
```

(Aside, `pull(mtcars, "cyl")` works as well)

When referring to an exsiting column inside a data frame, most (all?), tidyverse functions will not require quotes.  

## Strategy #2: Your Turn

The `gather()` function in tidyr takes multiple columns and collapses then into `key` and `value` columns.

Do the values of the `key` and `value` arguments need quotes?

```{r tidy=FALSE, printr.help.sections=c('arguments'), comment=''}
?tidyr::gather
```

## Strategy #2: Your Turn Solution

Do the values of the `key` and `value` arguments need quotes?

"as strings" -> YES!
```{r, results = "hide"}
gather(table4a, -country, key = "year", value = "count")
```

"or symbols" -> NO!
```{r, results = "hide"}
gather(table4a, -country, key =  year , value =  count )
```

So, in this case it doesn't matter.

## Strategy #3: `<-` Strategy #0

Reading documentation isn't always enlightening:

```{r eval = FALSE, tidy=FALSE, printr.help.sections=c('arguments'), comment=''}
?purrr::pluck
```
For the `...` argument:

> Can be an integer position, a string name, or an accessor function. 

A "string name", is that the name of a string, or name as a string?

Revert to strategy #0...

