---
title: Quotes, Quotation and Quasiquotation
author: Charlotte Wickham
date: May 8th 2018
output: github_document
---

# The game plan

Three examples:

1. Using a dplyr function interactively
2. Using a dplyr function with a saved variable
3. Using a dplyr function inside a function

```{r, message=FALSE}
library(tidyverse)
library(rlang)
```

Motivated by an [Exercise in Advanced R](https://adv-r.hadley.nz/quasiquotation.html#exercises-60)

> Implement `arrange_desc()`, a variant of `dplyr::arrange()` that sorts in descending order by default.

# 1. Using a dplyr function interactively

## dplyr

dplyr provides **data manipulation** verbs:
  
* `filter()`
* `select()`
* `arrange()`
* `mutate()`
* `summarise()`
* `group_by()`

## Star Wars

```{r}
starwars
```

## `arrange()`: Reorder rows in data

First argument, `.data`, is a data frame or tibble,  
other arguments, `...`, are columns to order by.  

Reorder by `mass`:
```{r}
arrange(starwars, mass)
```


Reorder by `mass`, breaking ties with `height`:
```{r}
arrange(starwars, mass, height)
```

Reorder by decreasing `mass`:
```{r}
arrange(starwars, desc(mass))
```

## The alternative with `order()` and `[`

```{r}
starwars[order(starwars$mass, starwars$height), ]
```

## Why dplyr?

* First argument is always a data frame / tibble, **amenable to piping**.
* Other arguments are evaluated in the context of the data, **saves typing**.
* Function names describe the action, **code is easier to read**.

```{r}
starwars %>%
  filter(height < 160) %>%
  select(name, height, mass) %>%
  arrange(mass)
```

## Quoted versus evaluated arguments

`order()` uses **standard evaluation**, its arguments are **evaluated**.  

That is, we can run the argument alone, it works, and it is all `order()` needs to do its job: 
```{r}
starwars$mass
```

`arrange()` uses **non-standard evaluation**, some of its arguments are **quoted**.   We can't just run the argument value alone, it doesn't work:
```{r, error = TRUE}
mass
```

## Why does dplyr use quotation?

dplyr uses quotation:

  * to save you typing! 
  * so dbplyr can translate to SQL

When you work interactively, this is great!

But, sometimes it causes problems...

# 2. Using a dplyr function with a saved variable

## Using `arrange()` with a saved variable

```{r}
arrange(starwars, desc(mass))
```

What if if the column to order by is stored in a variable?

This doesn't work:
```{r, error = TRUE}
col <- mass
arrange(starwars, col)
```

Neither does this:
```{r, error = TRUE}
col <- "mass"
arrange(starwars, col)
```

## A similar problem with `$`

`$` can be used to pull out a column from a data frame:
```{r}
starwars$mass
```

Even though you can't just ask for `mass`:
```{r, error = TRUE}
mass
```

So, `$` quotes its argument too.

But, what if the column I want is stored in a string?
```{r}
col <- "mass"
```

This doesn't work:
```{r, error = TRUE}
starwars$col
```

What to do? Use a function that doesn't use quote its argument:
```{r}
starwars[, col]
```

This is an example where you don't want the function to quote what you give it.  The solution in this case is to use a different function, one that doesn't quote its arguments.

The tidyverse implements a different solution: tidy evaluation.

## Tidy evaluation

An opinion on how **non-standard evaluation** should be implemented, and a set of tools for implementing it: the rlang package.

The way non-standard evaluation _is_ (will be) implemented in the tidyverse.

Two audiences:

1. For users of the tidyverse: tidy evaluation provides a way to selectively evaluate things if needed

2. For developers of tidy tools: tidy evaluation provides a framework for implementing your own non-standard evaluation tools, so your users get the benefits of #1.

We'll focus on the first case. We need two pieces: a way to quote things, and a way to selectively unquote things.

## #1: Quoting with rlang

Quoting means: to capture the intent of the code, not the result of evaluating it.

How do I capture the **intent** of this code:
`x + y + z`

If I just run it in R,
```{r, error = TRUE}
x + y + z
```
R tries to evaluate it, and I get an error, because some of the objects involved don't exist.

I could put quotes around it, to capture it without evaluating it:
```{r}
"x + y + z"
```
But this doesn't convey that this is in fact R code.

In rlang, one way is to use `expr()`:
```{r}
expr(x + y + z)
```
It returns a quoted expression.

(Base R has tools for quoting too, but rlang's implementation is more consistent)

## #2: Unquoting with `!!` (bang-bang)

**Inside** an rlang quoting function (e.g. `expr()`, and dplyr's quoted arguments), you can unquote something with the `!!` operator.

This quotes the expression `mass`:
```{r}
my_var <- expr(mass)
```

This quotes the expression `my_var`:
```{r}
expr(my_var)
```

This unquotes `my_var` before quoting the result:
```{r}
expr(!!my_var)
```

## Practice

```{r}
x <- expr(z)
y <- expr(x + y)
```

Can you guess what expression each of these will return?

```{r, results="hide"}
expr(x + y)
expr(!!x + y)
expr(x + !!y)
expr(!!x + !!y)
```

## Back to our problem

This didn't work:
```{r, error = TRUE}
col <- mass
arrange(starwars, col)
```

## A Solution

First we want to quote `height`, not evaluate it:
```{r}
var <- expr(mass)
```

Then we can unquote `var` when we pass it to `arrange()`:
```{r}
arrange(starwars, !!var)
```

You can selectively unquote things, so to get reverse ordering:
```{r}
arrange(starwars, desc(!!var))
```

# 3. Using a dplyr function inside a function

## Now, we might want to turn this into a function

```{r}
var <- expr(height)
arrange(starwars, desc(!!var))
```

One attempt:
```{r}
arrange_desc <- function(.data, var){
  arrange(.data, desc(!!var))
}

arrange_desc(starwars, expr(mass))
```
But it would be nicer if we could just say `arrange_desc(starwars, mass)`

## Another attempt

Why doesn't this work?
```{r, error = TRUE}
arrange_desc <- function(.data, var){
  var <- expr(var)
  arrange(.data, desc(!!var))
}
arrange_desc(starwars, mass)
```


```{r, eval = FALSE}
debugonce(arrange_desc)
arrange_desc(starwars, mass)
```

`expr(var)` returns the expression `var`, it gets quoted.  In rlang this is solved by using `enexpr()` instead.

## Use `enexpr()` instead of `expr()` to capture arguments inside of functions

And finally it works
```{r}
arrange_desc <- function(.data, var){
  var <- enexpr(var)
  arrange(.data, desc(!!var))
}
arrange_desc(starwars, mass)
```

But it only accepts one argument...

## A multiple argument version

New things:

* `exprs()` to capture multiple expressions in a list
* `purrr::map()` to operate of the list of expressions
* `!!!` to "splice" unquote: each element is inserted as an argument

```{r}
arrange_desc2 <- function(.data, ...){
  vars <- enexprs(...)
  vars_desc <- map(vars, function(var) expr(desc(!!var)))
  arrange(.data, !!!vars_desc)
}
arrange_desc2(starwars, mass, height)
```

## Quosures

In reality you'll see `quo()`, `enquo()`, `quos()` and `enquos()`, instead of `expr()`, `enexpr()`, `exprs()` and `enexprs()`.

These capture both the code and the **environment**.

# Resources 

If dplyr is new to you, start by learning to use it interactively with the [Data Transformation chapter in R for Data Science](http://r4ds.had.co.nz/transform.html).

If you have never written your own function before, start with [Functions in R for Data Science](http://r4ds.had.co.nz/functions.html) or [DataCamp's Writing Functions in R](https://www.datacamp.com/courses/writing-functions-in-r).

If you've written functions but want to formalise your knowledge, [Functions in Advanced R](http://adv-r.had.co.nz/Functions.html).

If you want to see how to program with dplyr functions, the [Programming with dplyr vignette](https://dplyr.tidyverse.org/articles/programming.html).

If the vignette seems a little unapproachable, find another resource in [Mara Averick's roundup of tidy eval resources](https://maraaverick.rbind.io/2017/08/tidyeval-resource-roundup/).

If you'd prefer to watch than read, the [RStudio's tidy evaluation webinar](https://www.rstudio.com/resources/webinars/tidy-eval/), or [Hadley's tidy evaluation in 5 minutes](https://www.youtube.com/watch?v=nERXS3ssntw).

If you are interested in the underpinnings, and possibly using tidy evaluation ideas in your own pacakges, look at [Metaprogramming in the **new** (in progress) edition of Advanced R](https://adv-r.hadley.nz/meta.html).


